# Nebulosa plots

This plot comes straight from the [Nebulosa package](http://bioconductor.org/packages/release/bioc/html/Nebulosa.html). For more information, here is their [vignette](http://bioconductor.org/packages/release/bioc/vignettes/Nebulosa/inst/doc/nebulosa_seurat.html). In short, it computes a "density" of whether the surrounding cells (in the UMAP embedding) also express the variable that you provide to the function, same variables that one could feed to `Seurat::FeaturePlot()`. This is how their default plot looks like:

```{r, fig.cap= "Nebulosa plot"}
Nebulosa::plot_density(sample, "CD14")
```

This way, not only we do know which clusters are enriched in CD14, we also know the regions of it with the hightest density of cells expressing it. Minor modifications, on the line of the rest of plots, are applied in `SCpubr`:
```{r, fig.cap = "SCpubr Nebulosa plot."}
SCpubr::do_NebulosaPlot(sample, features = "CD14")
```

Then, this type visualization becomes a natural partner to `Seurat::FeaturePlot()' as not only we are able to visualize the expression of a variable, but also query the density of the surrounding cells. Here is an example:

```{r, fig.cap = "SCpubr FeaturePlot and NebulosaPlot.", fig.height= 4, fig.width=8}
SCpubr::do_FeaturePlot(sample, features = "CD14") | SCpubr::do_NebulosaPlot(sample, features = "CD14")
```
